Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-23 parallel-tractability-J-v9/conclusions.tex
fixed some points in abstract, intro, related work, conclusion

In this paper, we studied the problem of parallel tractability of
materialization on the datalog rewritable ontologies.
To identify the parallel tractability classes,
we proposed two \texttt{NC} algorithms, Algorithm~$\mathsf{A}_{\text{bsc}}^\psi$
and Algorithm~$\mathsf{A}_{\text{opt}}^\psi$, that perform materialization on
datalog rewritable ontology languages.
Based on these algorithms, we identified the corresponding
parallel tractability classes such that materialization
on the datalog programs in these classes is in the NC complexity.
We further studied two specific ontology languages, DL-Lite and DHL (including one of its extension).
We showed that any ontology expressed in DL-Lite$_{core}$ or DL-Lite$_{\mathcal{R}}$ is tractable in parallel.
For DHL and DHL($\circ$), we proposed two restrictions such that materialization is tractable in parallel.

We analyzed the benchmark LUBM and the real-world dataset YAGO
and gave the cases where these ontologies belong to parallel tractability classes.
On the other hand, we used an optimization strategy based on SWD paths
to give a practical algorithm variant Algorithm~$\mathsf{A}_{\text{prc}}$, which can
also be restricted to an \texttt{NC} version.
We implemented a system based on Algorithm~$\mathsf{A}_{\text{prc}}$ and compared it to
the state-of-the-art reasoner RDFox.
The experimental results showed that our system is comparable to RDFox when handling
the LUBM and the YAGO ontologies.

In the future work, we will extend our work in two lines.
One line is a further study of the parallel tractability of
other expressive ontology languages, in addition to
datalog rewritable ones. Since different expressive OWL languages
have different syntaxes and higher reasoning complexities than \texttt{PTime}-complete, we need to
explore more restrictions that are practical and make materialization
tractable in parallel. Another line of the future work is to
study how to further apply the
results in practice.
One idea is to apply the technique of SWD paths to enhance other reasoning-based tasks,
like ontology classification, ontology debugging and query answering.

>>>>

In this paper, we studied the problem of parallel tractability of
materialization on the datalog rewritable ontologies.
To identify the parallel tractability classes,
we proposed two \texttt{NC} algorithms, Algorithm~$\mathsf{A}_{\text{bsc}}^\psi$
and Algorithm~$\mathsf{A}_{\text{opt}}^\psi$, that perform materialization on
datalog rewritable ontology languages.
Based on these algorithms, we identified the corresponding
parallel tractability classes such that materialization
on the datalog programs in these classes is in \texttt{NC} complexity.
We further studied two specific ontology languages, DL-Lite and DHL (including one of its extension).
We showed that any ontology expressed in DL-Lite$_{core}$ or DL-Lite$_{\mathcal{R}}$ is tractable in parallel.
For DHL and DHL($\circ$), we proposed two restrictions such that materialization is tractable in parallel.

We analyzed the benchmark LUBM and the real-world dataset YAGO
and gave the cases where these ontologies belong to parallel tractability classes.
On the other hand, we used an optimization strategy based on SWD paths
to give a practical algorithm variant Algorithm~$\mathsf{A}_{\text{prc}}$, which can
also be restricted to an \texttt{NC} version.
We implemented a system based on Algorithm~$\mathsf{A}_{\text{prc}}$ and compared it to
the state-of-the-art reasoner RDFox.
The experimental results showed that the optimizations proposed in this paper resulted in a
better performance on ontologies that were
tractable in parallel compared with RDFox.

In the future work, we will extend our work in two lines.
One line is a further study of the parallel tractability of
other expressive ontology languages, in addition to
datalog rewritable ontology languages. Since different expressive OWL languages
have different syntaxes and higher reasoning complexities than \texttt{PTime}-complete, we need to
explore more restrictions that are practical and make materialization
tractable in parallel. Another line of the future work is to
study how to further apply the
results in practice.
One idea is to apply the technique of SWD paths to enhance other reasoning-based tasks,
like ontology classification, ontology debugging and query answering.

### 2018-02-23 parallel-tractability-J-v9/introduction.tex
fixed some points in abstract, intro, related work, conclusion

In the practical part of this work, we study specific datalog rewritable ontology languages.
We first focus on the ontology language DL-Lite \cite{CalvaneseGLLR07} to clarify how to apply the above \texttt{NC} algorithms
to study parallel tractability of ontology materialization.
We show that DL-Lite$_{\text{core}}$ and DL-Lite$_\mathcal{R}$ are tractable in parallel.
We then study the ontology language Description Horn Logic (DHL) \cite{GrosofHVD03}, which is the intersection of datalog and OWL
in terms of expressivity. We give a case of a DHL ontology where materialization can hardly be parallelized.
Based on the analysis of this case, we propose to restrict the usage of DHL such that materialization over the
restricted ontologies can be handled by the proposed \texttt{NC} algorithms.
We further extend the results to an extension of DHL that also allows complex role inclusion axioms.
Finally, we analyze well-known benchmarks and real-world datasets and show that many real ontologies following the
proposed restrictions belong to the parallel tractability classes.
We implement a system based on an optimized \texttt{NC} algorithm for DHL materialization, which we compare to the state-of-the-art reasoner RDFox.
The experimental results show that the optimizations proposed in this paper result in a better performance on ontologies that are 
tractable in parallel compared with RDFox.

The remainder of the paper is organized as follows. In Section~\ref{sec:background}, we introduce some basic notions.
We then give two \texttt{NC} algorithms for ontology materialization in Section~\ref{sec:ptclass}.
We study parallel tractability of materialization in DL-Lite, DHL and an extension of DHL in Section~\ref{sec:ptonto}, respectively.
In Section~\ref{sec:practicalAlg},
we discuss how to optimize and implement the given \texttt{NC} algorithms.
We analyze real-world datasets and evaluate our implementation in Section~\ref{sec:evaluation}.
Finally, we discuss related work in Section~\ref{sec:related} and conclude in Section~\ref{sec:conclusion}. Detailed proofs of the theorems and lemmas in this paper are given in the appendix.

>>>>

In the practical part of this work, we study specific datalog rewritable ontology languages.
We first focus on the ontology language DL-Lite \cite{CalvaneseGLLR07} to clarify how to apply the above \texttt{NC} algorithms
to study parallel tractability of ontology materialization.
We show that DL-Lite$_{\text{core}}$ and DL-Lite$_\mathcal{R}$ are tractable in parallel.
We then study the ontology language Description Horn Logic (DHL) \cite{GrosofHVD03}, which is the intersection of datalog and OWL
in terms of expressivity. We give a case of a DHL ontology where materialization can hardly be parallelized.
Based on the analysis of this case, we propose to restrict the usage of DHL such that materialization over the
restricted ontologies can be handled by the proposed \texttt{NC} algorithms.
We further extend the results to an extension of DHL that also allows complex role inclusion axioms.
Finally, we analyze the well-known benchmark, LUBM, and the real-world dataset, YAGO, and
show that these ontologies following the
proposed restrictions belong to the parallel tractability classes.
We implement a system based on an optimized \texttt{NC} algorithm for DHL materialization,
which we compare to the state-of-the-art reasoner RDFox.
The experimental results show that the optimizations proposed in this paper result in a
better performance on ontologies that are
tractable in parallel compared with RDFox.

The remainder of the paper is organized as follows. In Section~\ref{sec:background}, we introduce some basic notions.
We then give two \texttt{NC} algorithms for ontology materialization in Section~\ref{sec:ptclass}.
We study parallel tractability of materialization in DL-Lite, DHL and an extension of DHL in Section~\ref{sec:ptonto}, respectively.
In Section~\ref{sec:practicalAlg},
we discuss how to optimize and implement the given \texttt{NC} algorithms.
We analyze the LUBM and the YAGO ontologies and evaluate our implementation in Section~\ref{sec:evaluation}.
Finally, we discuss related work in Section~\ref{sec:related} and conclude in Section~\ref{sec:conclusion}. Detailed proofs of the theorems and lemmas in this paper are given in the appendix.

### 2018-02-23 parallel-tractability-J-v9/parallel-tractability-J.tex
Proof-read eval section
fixed some points in abstract, intro, related work, conclusion

@@

\begin{abstract}
  Materialization is an important reasoning service for many
  ontology-based applications, but the rapid growth of semantic data
  poses the challenge to efficiently perform materialization on
  large-scale ontologies. Parallel materialization algorithms work
  well for some ontologies, although the reasoning problem for the
  used ontology language is not in \texttt{NC}, i.e., the theoretical
  complexity class for \emph{parallel tractability}. This motivates us
  to study the problem of \emph{parallel tractability} of ontology
  materialization from a theoretical perspective. We focus on the
  datalog rewritable ontology languages DL-Lite and Description Horn
  Logic (DHL) and propose algorithms, called \texttt{NC} algorithms,
  to identify classes of ontologies for which materialization is
  tractable in parallel. To verify the practical usability of the
  above results, we analyze benchmark and real-world datasets, including the LUBM
  and the YAGO ontologies, and show that for many
  ontologies expressed in DHL materialization is tractable in
  parallel. The implementation of our optimized parallel algorithm
  shows comparable performance with compared to the highly optimized
  state-of-the-art reasoner RDFox.
\end{abstract}

\begin{keyword}
ontology \sep materialization \sep datalog \sep parallel tractability \sep NC complexity
\end{keyword}

>>>>

\newcommand{\BG}[1]{\textcolor{red}{BG: #1}}
%\renewcommand{\BG}[1]{}

@@

\begin{abstract}
  Materialization is an important reasoning service for many
  ontology-based applications, but the rapid growth of semantic data
  poses the challenge to efficiently perform materialization on
  large-scale ontologies. Parallel materialization algorithms work
  well for some ontologies, although the reasoning problem for the
  used ontology language is not in \texttt{NC}, i.e., the theoretical
  complexity class for \emph{parallel tractability}. This motivates us
  to study the problem of \emph{parallel tractability} of ontology
  materialization from a theoretical perspective. We focus on the
  datalog rewritable ontology languages DL-Lite and Description Horn
  Logic (DHL) and propose algorithms, called \texttt{NC} algorithms,
  to identify classes of ontologies for which materialization is
  tractable in parallel. To verify the practical usability of the
  above results, we analyze different benchmarks and real-world datasets, including LUBM
  and the YAGO ontology, and show that for many
  ontologies expressed in DHL materialization is tractable in
  parallel. The implementation of our optimized parallel algorithm
  shows performance improvements over the highly optimized
  state-of-the-art reasoner RDFox on ontologies for which
  materialization is tractable in parallel.
\end{abstract}

\begin{keyword}
ontology \sep materialization \sep datalog \sep parallel tractability \sep \texttt{NC} complexity
\end{keyword}

### 2018-02-23 parallel-tractability-J-v9/relatedWork.tex
fixed some points in abstract, intro, related work, conclusion

The parallel reasoner RDFox \cite{MotikNPHO14} is used to evaluate our implementation. RDFox is a state-of-the-art system that handles reasoning on datalog rewritable ontology languages. Algorithm~$\mathsf{A}_{\text{bsc}}$ proposed in Section~\ref{sec:ptclass} is similar to the main algorithm of RDFox (see \cite{MotikNPHO14}, Sections~3 and 4). The difference lies in that, a group of rule instantiations are handled by one processor (namely a thread) in RDFox, while in Algorithm~$\mathsf{A}_{\text{bsc}}$, each rule instantiation is assigned to a unique processor. Analogous to the method used in RDFox, ParallelDHL is also implemented by assigning a group of
rule instantiations to a processor.
The experimental results show that ParallelDHL is comparable to RDFox when handling
the LUBM and the YAGO ontologies.

>>>>

The parallel reasoner RDFox \cite{MotikNPHO14} is used to evaluate our implementation. RDFox is a highly optimized system that handles reasoning on datalog rewritable ontology languages. Algorithm~$\mathsf{A}_{\text{bsc}}$ proposed in Section~\ref{sec:ptclass} is similar to the main algorithm of RDFox (see \cite{MotikNPHO14}, Sections~3 and 4). The difference lies in that, a group of rule instantiations are handled by one processor (namely a thread) in RDFox, while in Algorithm~$\mathsf{A}_{\text{bsc}}$, each rule instantiation is assigned to a unique processor.
Analogous to the method used in RDFox, ParallelDHL is also implemented by assigning a group of
rule instantiations to a processor.
Specially, ParallelDHL is also optimized by using SWD paths for handling
ontologies that belong to parallel tractability classes.
The experimental results show that this optimization results in a
better performance on ontologies that are
tractable in parallel compared with RDFox.
